---
title: "Homework: Queueing Theory"
author: Hayden Atchley
date: today
date-format: "D MMMM YYYY"
number-sections: false
execute: 
  echo: false
format: pdf
# include-in-header:
#   text: \usepackage[document]{ragged2e}
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(kableExtra)
library(queueing)

write_vec <- function(v){
  res <- case_when(
    length(v) < 1 ~ "NULL",
    length(v) == 1 ~ as.character(v),
    length(v) == 2 ~ paste(v[1], "and", v[2]),
    length(v) > 2 ~ paste(
      paste(v[1:(length(v)-1)], collapse = ", "),
      v[length(v)],
      sep = ", and "
    )
  )
  
  res[1]
}
```

<!-- \raggedright -->

# 1

A deterministic queueing theory is one in which constant behavior is assumed.
For example, an arrival rate of 60 individuals per hour in a deterministic theory means that each minute exactly one individual will arrive.
A stochastic theory, on the other hand, involves an element of randomness, i.e. there is a distribution which behavior follows. There will be e.g. an *average* arrival rate, but the exact arrival times are somewhat random.

# 2

```{r}
#| label: setup-2

arrival <- 100 #vph
service_time <- 20 #sec/veh
service_rate <- 3600/service_time #vph per station

wage <- 25 #$/hr
vot <- 10 #$/hr
```

A freeway on-ramp has a toll booth which takes `r service_time` seconds to service a vehicle, for a capacity rate of `r service_rate` vehicles per hour.
The peak period flow on this on-ramp is `r arrival` vehicles per hour.

```{r}
#| label: one-booth

mm1 <- NewInput.MM1(arrival, service_rate, n = 10) %>% 
  QueueingModel()

p_less_3 <- sum(mm1$Pn[1:4])
p_geq_3 <- 1 - p_less_3
```

This system is modeled as an M/M/1 Markov model, and solving this gives the following:
The average number of vehicles in the system (including any vehicles being served) is `r mm1$L` and the average wait time (before being served) is `r mm1$Wq %>% round(4)` hours or `r round(mm1$Wq*3600,1)` seconds.
Additionally, the probability of a queue 3 vehicles long or longer $P(\geq 4)$ (note that the queue does not include any vehicles currently being served) is given by $1 - \sum_{n=0}^3 P(n)$.
$P(0)$, $P(1)$, $P(2)$ and $P(3)$ are given by `r round(mm1$Pn[1:4],2) %>% write_vec()`, respectively, and so the probability of there being three or more vehicles in the queue is `r p_geq_3 %>% round(2)`.

```{r}
#| label: two-booths

mmc <- NewInput.MMC(arrival, service_rate, 2, n = 10) %>% 
  QueueingModel()

p_less_3 <- sum(mmc$Pn[1:4])
p_geq_3 <- 1 - p_less_3

wait_diff <- mm1$Wq - mmc$Wq
save_per_veh <- wait_diff*vot #$/veh
save_per_hour <- save_per_veh*arrival #$/hr
```

An additional toll booth is being considered, and this proposed scenario is modeled with an M/M/c (or M/M/N) Markov model where $c=2$ (or $N=2$).
Solving this model gives an individual expected wait time of `r (mmc$Wq*3600) %>% round(1)` seconds, which is `r (wait_diff*3600) %>% round(1)` seconds faster than the existing scenario.
At a regional value of time of \$`r vot`/hr, this saves `r round(save_per_veh*100,1)`&cent; per vehicle.
At an arrival rate of `r arrival` vehicles per hour, this saves \$`r round(save_per_hour,2)` per hour.
Since the toll operator wage is \$`r wage`/hr, this additional toll booth does not make much economic sense.

Additionally, the probability of three or more vehicles in the queue $P(\geq4)$ is calculated as above to give $P(\geq 4) = 1 - \sum_{n=0}^3{P(n)} = 1 - `r round(p_less_3,3)` = `r round(p_geq_3,3)`$.