---
title: "Homework: Queueing Theory"
author: Hayden Atchley
date: today
date-format: "D MMMM YYYY"
number-sections: false
execute: 
  echo: false
format: pdf
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(kableExtra)
library(queueing)

write_vec <- function(v){
  res <- case_when(
    length(v) < 1 ~ "NULL",
    length(v) == 1 ~ as.character(v),
    length(v) == 2 ~ paste(v[1], "and", v[2]),
    length(v) > 2 ~ paste(
      paste(v[1:(length(v)-1)], collapse = ", "),
      v[length(v)],
      sep = ", and "
    )
  )
  
  res[1]
}
```

# 1

A deterministic queueing theory is one in which constant behavior is assumed.
For example, an arrival rate of 60 individuals per hour in a deterministic theory means that each minute exactly one individual will arrive.
A stochastic theory, on the other hand, involves an element of randomness, i.e. there is a distribution which behavior follows. There will be e.g. an *average* arrival rate, but the exact arrival times are somewhat random.

# 2

```{r}
#| label: setup-2

arrival <- 100 #vph
service_time <- 20 #sec/veh
service_rate <- 3600/service_time #vph per station

wage <- 25 #$/hr
vot <- 10 #$/hr
```

A freeway on-ramp has a toll booth which takes `r service_time` seconds to service a vehicle, for a capacity rate of `r service_rate` vehicles per hour.
The peak period flow on this on-ramp is `r arrival` vehicles per hour.

```{r}
#| label: one-booth

mm1 <- NewInput.MM1(arrival, service_rate, n = 10) %>% 
  QueueingModel()

p_less_3 <- sum(mm1$Pn[1:4])
p_geq_3 <- 1 - p_less_3
```

Using the `queueing` R package gives the following values:
The average number of vehicles in the system (including any vehicles being served) is `r mm1$L` and the average wait time (before being served) is `r mm1$Wq %>% round(4)` hours or `r mm1$Wq*3600` seconds.
Additionally, the probability of a queue 3 vehicles long or longer $P(\geq 4)$ (note that the queue does not include any vehicles currently being served) is given by $1 - \sum_{n=0}^3 P(n)$.
$P(0)$, $P(1)$, $P(2)$ and $P(3)$ are given by `r round(mm1$Pn[1:4],2) %>% write_vec()`, respectively, and so the probability of there being three or more vehicles in the queue is `r p_geq_3 %>% round(2)`.