---
title: "Homework: Queueing Theory"
author: Hayden Atchley
date: today
date-format: "D MMMM YYYY"
number-sections: false
execute: 
  echo: false
format: pdf
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(kableExtra)
library(queueing)
```

# 1

A deterministic queueing theory is one in which constant behavior is assumed.
For example, an arrival rate of 60 individuals per hour in a deterministic theory means that each minute exactly one individual will arrive.
A stochastic theory, on the other hand, involves an element of randomness, i.e. there is a distribution which behavior follows. There will be e.g. an *average* arrival rate, but the exact arrival times are somewhat random.

# 2

```{r}
#| label: setup-2

arrival <- 100 #vph
service_time <- 20 #sec/veh
service_rate <- 3600/service_time #vph per station

wage <- 25 #$/hr
vot <- 10 #$/hr
```

A freeway on-ramp has a toll booth which takes `r service_time` seconds to service a vehicle, for a capacity rate of `r service_rate` vehicles per hour.
The peak period flow on this on-ramp is `r arrival` vehicles per hour.