---
title: "Homework: Equilibrium Traffic Flow"
subtitle: "4.1, 4.4, 4.8"
author: Hayden Atchley
date: today
date-format: "D MMMM YYYY"
number-sections: false
execute: 
  echo: false
format: pdf
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(flextable)
require(equatags)

hm <- function(vec){
  n <- length(vec)
  hm <- n / sum(1/vec)
  hm
}
```

# 4.1

Given a speed-density relationship of
\begin{align}
v = v_f\left(1-\frac{k}{k_j}\right), \label{vk}
\end{align}
and the relationship $q = k\times v$, we can derive flow-density and speed-flow relationships as follows:
\begin{align}
q &= kv \\
q &= kv_f\left(1-\frac{k}{k_j}\right) \\
q &= kv_f - \frac{k^2v_f}{k_j} \label{qk}
\end{align}  
and
\begin{align}
k = \frac{q}{v},\quad v &= v_f\left(1-\frac{k}{k_j}\right) \\
\implies v &= v_f\left(1-\frac{q/v}{k_j}\right) \\
\implies v^2 &= v_fv - \frac{qv_f}{k_j}. \label{vq}
\end{align}

From \eqref{qk} we can find $q_m$ (the capacity) and $k_m$ (the density at capacity) by determining the maximum of the flow-density relationship (where $\frac{dq}{dk} = 0$):
\begin{align}
q &= kv_f - \frac{k^2v_f}{k_j} \\
\frac{dq}{dk} = 0 &= v_f - 2\frac{k_mv_f}{k_j} \\
0 &= v_f\left(1 - 2\frac{k_m}{k_j}\right) \\
0 &= 1 - 2\frac{k_m}{k_j} \quad \text{if } v_f \neq 0 \\
k_m &= \frac{k_j}{2} \label{km}
\end{align}
\begin{align}
q_m &= k_mv_f \left(1 - \frac{k_m}{k_j}\right) \\
q_m &= \frac{k_jv_f}{2}\left(1 - \frac{k_j/2}{k_j}\right) \\
q_m &= \frac{k_jv_f}{4}. \label{qm}
\end{align}

$v_m$ (the speed at capacity) is then determined from \eqref{qm} and the original relationship \eqref{vk} by:
\begin{align}
v_m &= v_f\left(1 - \frac{k_m}{k_j}\right) \\
v_m &= v_f\left(1 - \frac{k_j/2}{k_j}\right) \\
v_m &= \frac{v_f}{2}.
\end{align}

{{< pagebreak >}}

# 4.4

The Greenberg model is given by
\begin{align}
v = v_m\ln{\frac{k_j}{k}}. \label{greenberg}
\end{align}

The capacity is again determined by setting $\frac{dq}{dk} = 0$, so a flow-density relationship is first determined:
\begin{align}
q = kv = kv_m(\ln k_j - \ln k). \label{gb-qk}
\end{align}

Then:
\begin{align}
\frac{dq}{dk} = 0 &= \frac{d}{dk} kv_m(\ln k_j - \ln k) \\
0 &= k_mv_m\left(-\frac{1}{k_m}\right) + v_m(\ln k_j - \ln k_m) \\
0 &= v_m\left(\ln \frac{k_j}{k_m} - 1\right) \\
1 &= \ln \frac{k_j}{k_m} \quad\text{if } v_m \neq 0 \\
e &= \frac{k_j}{k_m} \\
k_m &= \frac{k_j}{e}. \label{gb-km}
\end{align}

$q_m$ is then given from \eqref{gb-qk} and \eqref{gb-km}:
\begin{align}
q_m = k_mv_m(\ln k_j - \ln k_m) &= k_mv_m\left(\ln \frac{k_j}{k_m}\right) \\
q_m &= \frac{k_jv_m}{e}\left(\ln \frac{k_j}{k_j/e}\right) \\
q_m &= \frac{k_jv_m}{e}
\end{align}

{{< pagebreak >}}

# 4.8

```{r}
sk <- function(x){
  val <- 172 - 3.72*x + 0.0346*x^2 - 0.00119*x^3
  pmin(88.5, val)
}

qk <- function(x){
  sk(x)*x
}

kj <- uniroot(sk, c(0,200))$root
ff <- sk(0)

qopt <- optimize(qk, c(0,200), maximum = TRUE)

qm <- qopt$objective
km <- qopt$maximum
vm <- sk(km)

skpoints <- tribble(
  ~x, ~y, ~label,
  0, ff, paste("Free-flow speed =", round(ff,1)),
  kj, 0, paste("Jam density =", round(kj,1))
)

qkpoints <- tribble(
  ~x, ~y, ~label,
  km, qm, paste0("Capacity = ", round(qm), ", Density = ", round(km, 1))
)
```

The empirical speed-density relationship given by
\begin{align}
v = \min\{88.5,\ 172 - 3.72k + 0.0346k^2 - 0.00119k^3\} \label{empirical}
\end{align}
is shown graphically in @fig-4-8-vk.
The free-flow speed is the speed when the density is 0, or `r ff %>% round(1)` km/h.
The jam density $k_j$ is the upper bound of the valid range of the model (i.e. where the speed is nonnegative), which is `r kj %>% round(1)` veh/km.

```{r}
#| label: fig-4-8-vk
#| fig-cap: Empirical speed-density relationship.

ggplot(tibble(dens = c(0:floor(kj), kj)), aes(x = dens)) +
  stat_function(fun = sk) +
  labs(x = "Density (veh/km)", y = "Speed (km/h)") +
  scale_x_continuous(expand = expansion(mult = c(0,.05))) +
  scale_y_continuous(expand = expansion(mult = c(0,.05))) +
  theme_bw()
```

To find the capacity $q_m$ we need a flow-density plot.
Since $q = k\times v$, from \eqref{empirical} we get
\begin{align}
q = k\times \min\{88.5,\ 172 - 3.72k + 0.0346k^2 - 0.00119k^3\}.
\end{align}
@fig-4-8-qk shows this graphically.
$q_m$ is the maximum of this function, which is `r qm %>% round()` veh/h.
$k_m$ is the density at capacity ($q_m$), or `r km %>% round(1)` veh/km.
$v_m$ is the speed at this density, or `r vm %>% round(1)` km/h.

```{r}
#| label: fig-4-8-qk
#| fig-cap: Empirical flow-density relationship.

ggplot(tibble(dens = c(0:floor(kj), kj)), aes(x = dens)) +
  stat_function(fun = qk) +
  labs(x = "Density (veh/km)", y = "Flow (veh/h)") +
  scale_x_continuous(expand = expansion(mult = c(0,.05))) +
  scale_y_continuous(expand = expansion(mult = c(0,.05))) +
  theme_bw()
```

