---
title: "Homework: Conservation"
subtitle: "5.1, 5.3"
author: Hayden Atchley
date: today
date-format: "D MMMM YYYY"
number-sections: false
execute: 
  echo: false
format: pdf
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(flextable)
library(ftExtra)

set_flextable_defaults(
  float = "float",
)
```

# 5.1

```{r}
#| include: false

distance <- 1.5 #miles
interval <- 1 #hours
nlanes <- 2

tunnel <- read_csv("data/tunnel.csv") %>% 
  mutate(time = hour %>% hours() %>% as_datetime())

los <- tribble(
  ~los, ~smin, ~max,
  "A", -Inf, 11,
  "B", 11, 18,
  "C", 18, 26,
  "D", 26, 35,
  "E", 35, 45,
  "F", 45, Inf
) %>% 
  mutate(label = case_when(
    smin <= 0 ~ paste("&le;", max),
    max == Inf ~ paste("&gt;", smin),
    TRUE ~ paste0(smin, "&ndash;", max)
  ))

calc_tunnel <- tunnel %>% 
  mutate(
    sum_vin = cumsum(vin),
    sum_vout = cumsum(vout),
    vtunnel = sum_vin - sum_vout,
    dens = vtunnel / distance,
    dens_lane = dens / nlanes
    ) %>% 
  left_join(los, join_by(between(dens_lane, smin, max, bounds = "(]"))) %>% 
  select(-c(smin, max, label))
```

@tbl-tunnel shows vehicle counts from the Lincoln Tunnel's `r distance`-mile long south tunnel.
The counts are binned by hour over a 24-hour period, and are for both entering and exiting vehicles.

```{r}
#| label: tbl-tunnel
#| tbl-cap: Lincoln Tunnel Vehicle Counts

tunnel %>% 
  flextable(col_keys = c("time", "vin", "vout")) %>% 
  set_header_labels(time = "Time", vin = "Vehicles\nEntering", vout = "Vehicles\nExiting") %>% 
  colformat_datetime(fmt_datetime = "%H:%M") %>% 
  autofit() %>% 
  align(align = "c", part = "a")
```

The level of service (LOS) of the tunnel is determined by the density, as shown in @tbl-los.

```{r}
#| label: tbl-los
#| tbl-cap: LOS of Highway Segment

los %>% 
  flextable(col_keys = c("los", "label")) %>% 
  set_header_labels(los = "LOS", label = "Density (veh/mi/lane)") %>% 
  colformat_md() %>% 
  autofit() %>% 
  align(align = "c", part = "a")
```

For each hour, the density is calculated from the number of vehicles in the tunnel.
This is given by the difference in cumulative entering vehicles and cumulative exiting vehicles.
The tunnel is `r distance` miles in length and has `r nlanes` lanes, so the density per lane is given by $k_l = \frac{N_\text{tunnel}}{`r distance`\times `r nlanes`}$.
@tbl-calc-tunnel shows this for each hour.

```{r}
#| label: tbl-calc-tunnel
#| tbl-cap: Density and LOS Calculations for Lincoln Tunnel

calc_tunnel %>% 
  flextable(col_keys = c("time", "sum_vin", "sum_vout", "vtunnel", "dens", "dens_lane", "los")) %>% 
  set_header_labels(
    time = "Time", 
    sum_vin = "Cumulative\nVehicles Entering",
    sum_vout = "Cumulative\nVehicles Exiting",
    vtunnel = "Vehicles In&NewLine;Tunnel (En&minus;Ex)",
    dens = "Density\n(veh/mi)",
    dens_lane = "Lane Density\n(veh/mi/lane)",
    los = "Level of\nService"
  ) %>% 
  colformat_datetime(fmt_datetime = "%H:%M") %>% 
  colformat_double(j = ~ dens + dens_lane, digits = 2) %>% 
  colformat_md(j = vtunnel, part = "h") %>% 
  autofit() %>% 
  fit_to_width(7) %>% 
  align(align = "c", part = "a")
```

# 5.3



